---
title: "Data Sources"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Data_Sources}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

<style>
div.description { background-color:#FFCC33; border-radius: 20px; padding: 10px;}
</style>

Below is a record of the data products synthesised during MiMeMo. I've kept track of the structure of the data to make it easier to quickly check what we have. This will also function as a README for anyone else who follows and wants signposting to sources, or to know what additional data could be downloaded. 


# Ecological Drivers

## EMEP

<div class = "description"> Modelled atmospheric deposition of reduced and oxidised Nitrogen from EMEP MSC-W. </div>


#### <i class="fa fa-globe" aria-hidden="true" style ="color:#003399;"></i>  Spatial extent

Europe. 


#### <i class="fa fa-stopwatch" aria-hidden="true" style ="color:#003399;"></i>  Temporal extent

* *From* 2000/01/01. 
* *To* 2017/12/28.
* *In* months.
* *Duration* 18 years.


#### <i class="fa fa-folder" aria-hidden="true" style ="color:#003399;"></i>  File structure

* 18 netcdf files.
* 15gb total memory requirement.
* 1 netcdf file contains a 2D spatial grid and all months within a year.


#### <i class="fa fa-eye" aria-hidden="true" style ="color:#003399;"></i>  Target variables

* "DDEP_OXN_m2Grid", mean dry deposition of oxidised nitrogen.
* "WDEP_OXN", mean wet deposition of oxidised nitrogen.
* "DDEP_RDN_m2Grid", mean dry deposition of reduced nitrogen.
* "WDEP_RDN", mean wet deposition of reduced nitrogen.

#### <i class="fa fa-layer-group" aria-hidden="true" style ="color:#003399;"></i>  Spatial variables

*Located within netcdf dims, not as variables.*

* "lon", vector of latitudes (rectilinear grid).
* "lat", vector of longitudes (rectilinear grid).


#### <i class="fa fa-plus" aria-hidden="true" style ="color:#003399;"></i>  Other variables

"SURF_ugN_NOX", "SURF_ugN_RDN", "SURF_ugN_OXN", "WDEP_PREC", "WDEP_SOX", "SURF_MAXO3", "SURF_PM25water", "SOMO35", "DDEP_SOX_m2Grid", "DDEP_SOX_m2Conif", "DDEP_SOX_m2Decid", "DDEP_SOX_m2Seminat", "DDEP_OXN_m2Conif", "DDEP_OXN_m2Decid", "DDEP_OXN_m2Seminat", "DDEP_RDN_m2Conif", "DDEP_RDN_m2Decid", "DDEP_RDN_m2Seminat", "MMAOT40_IAM_DF", "MMAOT40_IAM_CR", "EUAOT40_Crops", "EUAOT40_Forests", "SURF_ug_SO2", "SURF_ug_NH3", "SURF_ug_HNO3", "SURF_ug_NO2", "SURF_ug_NO", "SURF_ug_SO4", "SURF_ug_NO3_F", "SURF_ug_NO3_C", "SURF_ug_NH4_F", "SURF_ppb_O3", "SURF_ug_PM10", "SURF_ug_PPM25", "SURF_ug_PPM_C", "SURF_ug_SS", "SURF_ug_DUST", "SURF_ug_ECFINE", "SURF_ug_ECCOARSE", "SURF_ug_PM25", "SURF_ug_PM25_rh50", "SURF_ug_PM10_rh50".
<br><br>


## EMCWF

<div class = "description"> Significant wave height and direction from the CERA-20C "Ocean Wave Synoptic Monthly Means". </div>


#### <i class="fa fa-globe" aria-hidden="true" style ="color:#003399;"></i>  Spatial extent

Northern hemisphere. 

*A global download is avaialable.*


#### <i class="fa fa-stopwatch" aria-hidden="true" style ="color:#003399;"></i>  Temporal extent

*Back to 1901 is avaiable as a daily time series if required.*

* *From* 2000/01/01. 
* *To* 2010/12/28.
* *In* 96 timesteps (every 3 hours for a synoptic day per month).
* *Duration* 11 years.


#### <i class="fa fa-folder" aria-hidden="true" style ="color:#003399;"></i>  File structure

* 11 netcdf files.
* 78.3gb total memory requirement.
* 1 netcdf file contains a 2D spatial grid and all timesteps within a year.


#### <i class="fa fa-eye" aria-hidden="true" style ="color:#003399;"></i>  Target variables

* "swh", Mean significant wave height.
* "mwd", Mean wave direction.
* "mwp", Mean wave period.


#### <i class="fa fa-layer-group" aria-hidden="true" style ="color:#003399;"></i>  Spatial variables

*Located within netcdf dims, not as variables.*

* "latitude", vector of latitudes (rectilinear grid).
* "longitude", vector of longitudes (rectilinear grid).


#### <i class="fa fa-plus" aria-hidden="true" style ="color:#003399;"></i>  Other variables

*Additional variables are available, but have not been downloaded from the web portal.*

"10 metre wind direction", "10 metre wind speed", "Air density over the oceans", "Benjamin-Feir index", "Coefficient of drag with waves", "Free convective velocity over the oceans", "Maximum individual wave height", "Mean direction of total swell", "Mean direction of wind waves", "Mean period of total swell", "Mean period of wind waves", "Mean square slope of waves", "Mean wave direction of first swell partition", "Mean wave direction of second swell partition", "Mean wave direction of third swell partition", "Mean wave period based on first moment", "Mean wave period based on first moment for swell", "Mean wave period based on first moment for wind waves", "Mean wave period based on second moment for swell", "Mean wave period based on second moment for wind waves", "Mean wave period of first swell partition", "Mean wave period of second swell partition", "Mean wave period of third swell partition", "Mean zero-crossing wave period", "Normalized energy flux into ocean", "Normalized energy flux into waves", "Normalized stress into ocean", "Peak wave period", "Period corresponding to maximum individual wave height", "Significant height of total swell", "Significant height of wind waves", "Significant wave height of first swell partition", "Significant wave height of second swell partition", "Significant wave height of third swell partition", "U-component stokes drift", "V-component stokes drift", "Wave Spectral Skewness", "Wave spectral directional width", "Wave spectral directional width for swell", "Wave spectral directional width for wind waves", "Wave spectral kurtosis", "Wave spectral peakedness".

<br><br>

## GEBCO

<div class = "description"> Global terrain model for ocean and land at 15 arc-second intervals. </div>

#### <i class="fa fa-globe" aria-hidden="true" style ="color:#003399;"></i>  Spatial extent

Global. 


#### <i class="fa fa-stopwatch" aria-hidden="true" style ="color:#003399;"></i>  Temporal extent

Snapshot, 2019.


#### <i class="fa fa-folder" aria-hidden="true" style ="color:#003399;"></i>  File structure

* 1 netcdf file.
* 11gb total memory requirement.
* The file contains a global 2D spatial grid.


#### <i class="fa fa-eye" aria-hidden="true" style ="color:#003399;"></i>  Target variables

* "elevation", bathymetry and topography.


#### <i class="fa fa-layer-group" aria-hidden="true" style ="color:#003399;"></i>  Spatial variables

* "lat", vector of latitudes (rectilinear grid).
* "lon", vector of longitudes (rectilinear grid).


#### <i class="fa fa-plus" aria-hidden="true" style ="color:#003399;"></i>  Other variables

NA

<br><br>


## Globcolour

<div class = "description"> Suspended particulate matter from binned monthly satellite colour data. </div>


#### <i class="fa fa-globe" aria-hidden="true" style ="color:#003399;"></i>  Spatial extent

Global. 


#### <i class="fa fa-stopwatch" aria-hidden="true" style ="color:#003399;"></i>  Temporal extent

* *From* 1997/09/01. 
* *To* 2020/31/28.
* *In* months.
* *Duration* 22 years.


#### <i class="fa fa-folder" aria-hidden="true" style ="color:#003399;"></i>  File structure

* 269 netcdf files.
* 17.5gb total memory requirement.
* 1 netcdf file contains a 2D spatial grid of mean values within a month.


#### <i class="fa fa-eye" aria-hidden="true" style ="color:#003399;"></i>  Target variables

* "SPM-0C5_mean", mean inorganic suspended particulate matter in sea water.


#### <i class="fa fa-layer-group" aria-hidden="true" style ="color:#003399;"></i>  Spatial variables

* "col", vector of longitudinal indices of the bins stored in the file.
* "center_lat", vector of centre latitudes for each useful row.
* "centerlon", vector of centre longitudes of the first left bin in a row.
* "lon_step", vector of longitudinal steps for each useful row.

Two dimensions are also required:

* "bin", id for a pixel
* "row", the row in the grid for a pixel.

To get to coordinates in latitude and longitude:

```{r code Globcolour coordinates, eval = FALSE, warning= FALSE}
  index <- row[bin] - row[1]+1 
  lat_bin = center_lat[index]
  lon_bin = center_lon[index] + col[bin] * lon_step[index]
```

#### <i class="fa fa-plus" aria-hidden="true" style ="color:#003399;"></i>  Other variables

*Additional variables are available, but have not been downloaded from the server.*

"CHL1", "CHL-OC5", "CHL2", "TSM", "PIC", "POC", "NFLH", "WVCS", "Txxx", "Axxx", "CF", "ABSD", "(N)RRSxxx", "PAR", "BBP", "CDM", "KD490", "KDPAR", "ZHL", "ZEU", "ZSD", "CHL-CIA", "BBPxxx-LOG", "BBPS-LOG", "PSD-XXX", "POC-SURF", "POC-INT", "PP", "PP-AM", "PP-UITZ", "PHYSAT".

<br><br>


## NEMO-MEDUSA

### Driving data

<div class = "description"> Driving data **FOR** NEMO-MEDUSA includes surface irradiance and air temperature at the sea surface. </div>

### Outputs

<div class = "description"> NEMO-MEDUSA model **outputs** extracted for StrathE2EPolar include:
Water temperature,
DIN,
Blah, </div>

## NGU

<div class = "description"> Sediment classes in the Barents Sea, an overview document of likely bottom types from expert synthesis of multiple data sets. </div>


#### <i class="fa fa-globe" aria-hidden="true" style ="color:#003399;"></i>  Spatial extent

Barents Sea. 


#### <i class="fa fa-stopwatch" aria-hidden="true" style ="color:#003399;"></i>  Temporal extent

Snapshot, 2015.


#### <i class="fa fa-folder" aria-hidden="true" style ="color:#003399;"></i>  File structure

* 1 shape file.
* 8mb total memory requirement.
* The file contains a collection of spatial polygons where sediment classes are expected to occur within the Barents Sea.


#### <i class="fa fa-eye" aria-hidden="true" style ="color:#003399;"></i>  Target variable

* "SEDKORNSTR", Sediment grain size code.
* "SEDKO_NAVN" , Sediment grain size class name.
 

#### <i class="fa fa-layer-group" aria-hidden="true" style ="color:#003399;"></i>  Spatial variables

Polygon boundaries embedded in shapefile.


#### <i class="fa fa-plus" aria-hidden="true" style ="color:#003399;"></i>  Other variables

NA

<br><br>


## SINMOD

<div class = "description"> Tidally resolved water velocity fields in the Arctic. Numerical ocean model output. </div>


#### <i class="fa fa-globe" aria-hidden="true" style ="color:#003399;"></i>  Spatial extent

Arctic. 


#### <i class="fa fa-stopwatch" aria-hidden="true" style ="color:#003399;"></i>  Temporal extent

* *From* 2003/01/01. 
* *To* 2018/12/28.
* *In* 336 timesteps (every 2 hours in a 28 day month).
* *Duration* 16 years.


#### <i class="fa fa-folder" aria-hidden="true" style ="color:#003399;"></i>  File structure

* 191 netcdf files.
* 600gb total memory requirement.
* 1 netcdf file contains a 3D spatial grid and all timesteps within a month.


#### <i class="fa fa-eye" aria-hidden="true" style ="color:#003399;"></i>  Target variables

* "u_east", Zonal veolcity.
* "v_north", Meridional velocity.


#### <i class="fa fa-layer-group" aria-hidden="true" style ="color:#003399;"></i>  Spatial variables

* "gridLats", matrix of latitudes (curvilinear grid).
* "gridLons", matrix of longitudes (curvilinear grid).
* "depths", matrix of depths.
* "ldepths", vector of depth layer thicknesses?


#### <i class="fa fa-plus" aria-hidden="true" style ="color:#003399;"></i>  Other variables

"w_east", "w_north", "w_velocity", "vertical_diffusivity", "NO3", "PhytoF", "PhytoD", "temperature",
"salinity", "ice_thickness", "ice_compactness", "freezing_rate", "salinity_ice", "uIce_east", "vIce_north", "uUnderIceE", "vUnderIceN", "elevation".

<br><br>

# Fishing Fleet

## Global Fishing Watch

<div class = "description"> Fishing effort from anonymised VMS signals and a model of vessel movement patterns. </div>


#### <i class="fa fa-globe" aria-hidden="true" style ="color:#003399;"></i>  Spatial extent

Global. 


#### <i class="fa fa-stopwatch" aria-hidden="true" style ="color:#003399;"></i>  Temporal extent

* *From* 2012/01/01. 
* *To* 2016/12/31.
* *In* days.
* *Duration* 5 years.


#### <i class="fa fa-folder" aria-hidden="true" style ="color:#003399;"></i>  File structure

* 1,827 csv files.
* 18gb total memory requirement.
* 1 csv file contains a list of pixels from the global grid where fishing was detected for a day.


#### <i class="fa fa-eye" aria-hidden="true" style ="color:#003399;"></i>  Target variables

* "geartype", fishing gear used.
* "fishing_hours", total hours of fishing within a pixel for a given day).


#### <i class="fa fa-layer-group" aria-hidden="true" style ="color:#003399;"></i>  Spatial variables

* "lat_bin", latitude * 100  (rectilinear grid).
* "lon_bin", longitude * 100  (rectilinear grid).


#### <i class="fa fa-plus" aria-hidden="true" style ="color:#003399;"></i>  Other variables

"flag", "vessel_hours", "mmsi_present".

<br>

## ICES

<div class = "description"> Fisheries landings by species and country. </div>


#### <i class="fa fa-globe" aria-hidden="true" style ="color:#003399;"></i>  Spatial extent

Major fishing area 27. 


#### <i class="fa fa-stopwatch" aria-hidden="true" style ="color:#003399;"></i>  Temporal extent

* *From* 1903/01/01. 
* *To* 2017/12/31.
* *In* years.
* *Duration* 115 years.


#### <i class="fa fa-folder" aria-hidden="true" style ="color:#003399;"></i>  File structure

* 1 csv file.
* 37mb total memory requirement.
* The file contains a list of landings by species, FAO areas, and flag for the whole time span.


#### <i class="fa fa-eye" aria-hidden="true" style ="color:#003399;"></i>  Target variables

* "tonnage", weight of fish landed.
* "year", year of lannding.
* "cname", country name.
* "N.blocs", block of nations.
* "Category", species guild.
* "Subcategory", species guild.
* "scientific", species name.


#### <i class="fa fa-layer-group" aria-hidden="true" style ="color:#003399;"></i>  Spatial variables

* "area", FAO region.
* "area_orig", old FAO naming convention.


#### <i class="fa fa-plus" aria-hidden="true" style ="color:#003399;"></i>  Other variables

"ccode2", "ccode3", acronym", "tsn", "common".

<br>


# Target data

## Blanchard report
